Followers Are Not Enough: Beyond Structural Communities in Online Social Networks

### 2014-04-30 paper-for_arxiv/methods.tex
Minor revisions.

Let $\{X_{t}\}$ and $\{ Y_{t}\}$ be two strong-sense stationary stochastic processes\footnote{Recall that a a stochastic process is strong-sense stationary if the distribution over process evaluated at finitely many time points is invariant to an overall timeshift~\cite{grimmett2001probability}.}. In our work, these would correspond to two user activities, $X_{t}(u)$ and $X_{t}(v).$ We use the notation $X_{t-k}^{t}$ to denote the values of the stochastic process from time $t-k$ to time $t$, $X_{t-k}^{t} = (X_{t-k}, X_{t-(k-1)}, \ldots, X_{t - 1}, X_{t})$. The lag-$k$ transfer entropy of $Y$ on $X$ is defined as 
\begin{align}
	\text{TE}_{Y \to X}^{(k)} &= H\left[X_{t} | X_{t-k}^{t-1}\right] - H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right],
\end{align}
where
\begin{align}
	H\left[X_{t} | X_{t-k}^{t-1}\right] = - E\left[ \log_{2} p(X_{t} | X_{t-k}^{t-1}) \right]
\end{align}
and 
\begin{align}
	H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right] = - E\left[ \log_{2} p(X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}) \right]
\end{align}
are the usual conditional entropies over the conditional (predictive) distributions $p(x_{t} | x_{t-k}^{t-1})$ and $p(x_{t} | x_{t-k}^{t-1}, y_{t-k}^{t-1})$. This formulation was originally developed in~\cite{schreiber2000measuring}, where transfer entropy was proposed as an information theoretic measure of \emph{directed} information flow. Formally, recalling that $H\left[X_{t} | X_{t-k}^{t-1}\right]$ is the uncertainty in $X_{t}$ given its values at the previous $k$ time points, and that $H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right]$ is the uncertainty in $X_{t}$ given the joint process $\{(X_{t}, Y_{t})\}$ at the previous $k$ time points, transfer entropy measures the reduction in uncertainty of $X_{t}$ by including information about $Y_{t-k}^{t-1},$ controlling for the information in $X_{t - k}^{t-1}$. By the `conditioning reduces entropy' result~\cite{cover2012elements}
\begin{align}
	H[X | Y, Z] \leq H[X | Y],
\end{align}
we can see that transfer entropy is always non-negative, and is zero precisely when $H\left[X_{t} | X_{t-k}^{t-1}\right] = H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right]$, in which case knowing the past $k$ lags of $Y_{t}$ does not reduce the uncertainty in $X_{t}$. If the transfer entropy is positive, then $\{ Y_{t}\}$ is considered causal for $\{ X_{t}\}$ in the Granger sense~\cite{granger1963economic}.
% More generally we can define the transfer entropy as the limit of lag-$k$ transfer entropies,
% \begin{align}
% 	\text{TE}_{Y \to X} &= \lim_{k \to \infty} \text{TE}_{Y \to X}^{(k)},
% \end{align}
% if the limit exists.

>>>>

Let $\{X_{t}\}$ and $\{ Y_{t}\}$ be two strong-sense stationary stochastic processes\footnote{Recall that a stochastic process is strong-sense stationary if the joint distribution for the process evaluated at finitely many time points is invariant to an overall timeshift~\cite{grimmett2001probability}.}. In our work, these would correspond to the activities, $X_{t}(u)$ and $X_{t}(v),$ of two users $u$ and $v$. We use the notation $X_{t-k}^{t}$ to denote the values of the stochastic process from time $t-k$ to time $t$, $X_{t-k}^{t} = (X_{t-k}, X_{t-(k-1)}, \ldots, X_{t - 1}, X_{t})$. The lag-$k$ transfer entropy of $Y$ on $X$ is defined as 
\begin{align}
	\text{TE}_{Y \to X}^{(k)} &= H\left[X_{t} | X_{t-k}^{t-1}\right] - H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right],
\end{align}
where
\begin{align}
	H\left[X_{t} | X_{t-k}^{t-1}\right] = - E\left[ \log_{2} p(X_{t} | X_{t-k}^{t-1}) \right]
\end{align}
and 
\begin{align}
	H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right] = - E\left[ \log_{2} p(X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}) \right]
\end{align}
are the usual conditional entropies over the conditional (predictive) distributions $p(x_{t} | x_{t-k}^{t-1})$ and $p(x_{t} | x_{t-k}^{t-1}, y_{t-k}^{t-1})$. This formulation was originally developed in~\cite{schreiber2000measuring}, where transfer entropy was proposed as an information theoretic measure of \emph{directed} information flow. Formally, recalling that $H\left[X_{t} | X_{t-k}^{t-1}\right]$ is the uncertainty in $X_{t}$ given its values at the previous $k$ time points, and that $H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right]$ is the uncertainty in $X_{t}$ given the joint process $\{(X_{t}, Y_{t})\}$ at the previous $k$ time points, transfer entropy measures the reduction in uncertainty of $X_{t}$ by including information about $Y_{t-k}^{t-1},$ controlling for the information in $X_{t - k}^{t-1}$. By the `conditioning reduces entropy' result~\cite{cover2012elements}
\begin{align}
	H[X | Y, Z] \leq H[X | Y],
\end{align}
we can see that transfer entropy is always non-negative, and is zero precisely when $H\left[X_{t} | X_{t-k}^{t-1}\right] = H\left[X_{t} | X_{t-k}^{t-1}, Y_{t-k}^{t-1}\right]$, in which case knowing the past $k$ lags of $Y_{t}$ does not reduce the uncertainty in $X_{t}$. If the transfer entropy is positive, then $\{ Y_{t}\}$ is considered causal for $\{ X_{t}\}$ in the Granger sense~\cite{granger1963economic}.
% More generally we can define the transfer entropy as the limit of lag-$k$ transfer entropies,
% \begin{align}
% 	\text{TE}_{Y \to X} &= \lim_{k \to \infty} \text{TE}_{Y \to X}^{(k)},
% \end{align}
% if the limit exists.
